---
title: "Movement matrix goodness of fit"
author: "Benjamin Galuardi"
date: today
format:
  html:
    toc: true
    embed-resources: true
    code-fold: show
execute:
  echo: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| echo: true
#| eval: true
library(SatTagSim)
library(knitr)
library(truncnorm)
library(fields)

data(sim_example)
data(box7)

mycol = colorRampPalette(c("lightcyan", "royalblue", "blue", "lemonchiffon", "orange", "red"), space = "Lab")
```

#### Generate seasonal transition matrices

Using the pre-simulated dataset (`simdat`) and the 7-box spatial stratification (`box7`):

```{r}
#| label: get_trans_prob
#| echo: true
#| eval: true
datbox <- get.first.box(simdat, 1000, box7, seas.len = 24)
trans.ct <- get.trans.prob(datbox, nyears = 50, adims = c(7, 7, 4), perc = FALSE)
trans <- get.trans.prob(datbox, nyears = 50, adims = c(7, 7, 4), perc = TRUE)
seasons <- c("Winter", "Spring", "Summer", "Fall")
names(trans.ct) <- names(trans) <- seasons
```

### Functions for multinomial variance

The expected number of times outcome $i$ is observed over $n$ trials is:
\begin{equation}
{\displaystyle \operatorname {E} (X_{i})=np_{i}\,} 
\end{equation}

\begin{equation}
\operatorname{v}(X_i) = np_i(1-p_i)
\end{equation}

where $n$ is the number of trials and $p$ is the probability. Here, each row of a movement matrix is considered a multinomial.

```{r}
#| label: multinomial_variance
#| echo: true
#| eval: true
# function to calculate variance of a multinomial row
get.mvar <- function(x){
  p <- drop(x) # use percentages
  p * (1 - p)
}

# function to sample from a posterior multinomial distribution
get.mvar.samp <- function(vmu, vsd) {
  o <- truncnorm::rtruncnorm(1, mean = vmu, sd = sqrt(vsd), a = 0, b = 1)
  o[is.nan(o)] <- 0
  o
}

# calculate variance matrices across seasons
vmat <- lapply(trans, function(x) apply(x, 1, get.mvar))
```

```{r}
#| label: vmattab
#| echo: true
#| eval: true
kable(vmat[[1]], col.names = 1:7, caption = "Multinomial variance for Winter", digits = 4)
```

{{< pagebreak >}}

## Plot the variance for each cell, each season
### Overlay with original movement matrix values

```{r}
#| label: plot_variance
#| echo: true
#| eval: true
#| fig-width: 9
#| fig-height: 9
#| results: hide
ex7 <- expand.grid(1:7, 1:7)

par(mfrow = c(2, 2))
sapply(1:4, function(x) {
  fields::image.plot(1:7, 1:7, vmat[[x]], col = mycol(49), xlab = '', ylab = '')
  text(ex7[, 1], ex7[, 2], round(as.vector(trans[[x]]), 3), font = 2)
  title(names(trans)[x])
})
```

----

{{< pagebreak >}}

## Sample from posterior
```{r}
#| label: sample_posterior
#| echo: true
#| eval: true
samp <- data.frame(sapply(1:7, function(x) get.mvar.samp(trans[[1]][x, ], vmat[[1]][x, ])))
names(samp) <- 1:7
rownames(samp) <- 1:7
kable(samp, caption = 'Posterior sample: Winter season', digits = 4)
```

----

{{< pagebreak >}}

## Goodness of Fit

### Functions to compare pairs of rows using log-linear models and $\chi^2$ test

$\chi^{2}$ multinomial goodness of fit:

\begin{equation}
{\displaystyle \operatorname {\chi^{2}} = \sum_{i} \frac{(f_{i}-e_{i})^2}{e_{i}}\,} 
\end{equation}

```{r}
#| label: gof_functions
#| echo: true
#| eval: true
# first argument is counts, second argument is probability.
# Row-wise comparisons of two matrices: p-values of log-linear (chisq test) are returned

get.loglin <- function(mat1, mat2){
  fm <- list()
  for(i in 1:nrow(mat1)){
    fm[[i]] <- loglin(rbind(mat1[i, ], mat2[i, ] + 1e-30), c(1), print = FALSE, fit = FALSE)
  }
  unlist(lapply(fm, function(x) 1 - pchisq(x$lrt, x$df)))
}

get.mgof <- function(mat1, mat2){
  fm <- numeric(nrow(mat1))
  for(i in 1:nrow(mat1)){
    if (sum(mat1[i, ]) == 0) {
      fm[i] <- NA
    } else {
      p <- mat2[i, ] + 1e-30
      p <- p / sum(p)
      fm[i] <- suppressWarnings(chisq.test(mat1[i, ], p = p)$p.value)
    }
  }
 fm
}
```

Since each row of the movement matrix depends on its associated region in the previous time step, each row is evaluated as its own 1-dimensional contingency table.

{{< pagebreak >}}

### Compare between seasons (e.g. Winter vs Spring)

```{r}
#| label: compare_seasons
#| echo: true
#| eval: true
bw.bs <- get.mgof(trans.ct[[1]], trans[[2]])
print(bw.bs)
```

### Compare consecutive seasons

```{r}
#| label: compare_consecutive_seasons
#| echo: true
#| eval: true
seas_comp <- data.frame(
  "Winter_vs_Spring" = get.mgof(trans.ct[[1]], trans[[2]]),
  "Spring_vs_Summer" = get.mgof(trans.ct[[2]], trans[[3]]),
  "Summer_vs_Fall"   = get.mgof(trans.ct[[3]], trans[[4]]),
  "Fall_vs_Winter"   = get.mgof(trans.ct[[4]], trans[[1]])
)
rownames(seas_comp) <- 1:7
kable(seas_comp, caption = "Multinomial goodness of fit (p-values) between consecutive seasons", digits = 4)
```

{{< pagebreak >}}

### Overall transition rate across 4 seasons

```{r}
#| label: annual_transition
#| echo: true
#| eval: true
# annual transition rate obtained by multiplying seasonal matrices
ball <- trans[[1]] %*% trans[[2]] %*% trans[[3]] %*% trans[[4]]
round(ball, 3)
```

----

{{< pagebreak >}}

### Tradeoff between degrees of freedom and likelihood ratio within the $\chi^{2}$ test

```{r}
#| label: pchisq_plot
#| echo: true
#| eval: true
plot(pchisq(rep(50, 100), df = 1:100), typ = 'l', lwd = 2, col = 4, xlab = 'DF', ylab = 'Value')
title('pchisq value')
```
