---
title: "Movement matrix goodness of fit"
author: "Benjamin Galuardi"
date: '`r Sys.Date()`'
output: pdf_document
header-includes: \usepackage{color, enumitem}
---

```{r, echo = F, eval = T, message = F, warning  = F}
library(SatTagSim)
library(knitr)

username =  as.character(Sys.info()[7])
setwd(paste0('C:/Users/', username,'/Google Drive/PHD/COLLABORATIONS/SCRS_SIMS/'))
datadir = 'D:/SKYDRIVE_BACKUP/PHD_SIMS/SCRS_SIMS'

mycol = colorRampPalette(c("lightcyan", "royalblue", "blue", "lemonchiffon", "orange", "red"), space = "Lab")
load('sim_comparison_models.Rdata')
```

#### Variables used in this example (three datasets/runs)
 + East Atlantic
 + West Atlantic large fish
 + West Atlantic small fish

```{r, echo = T, eval = F}

eatl.ct = get.trans.prob(eatl.datbox, perc = F)
watl.big.ct = get.trans.prob(watl.big.datbox, perc = F)
watl.small.ct = get.trans.prob(watl.small.datbox, perc = F)
eatl = get.trans.prob(eatl.datbox, perc = T)
watl.big = get.trans.prob(watl.big.datbox, perc = T)
watl.small = get.trans.prob(watl.small.datbox, perc = T)

```

### functions for multinomial variance. 

The expected number of times the outcome $i$ observed over $n$ trials is
\begin{equation}
{\displaystyle \operatorname {E} (X_{i})=np_{i}\,} 
\end{equation}

\begin{equation}
\operatorname{v}(X_i) = np_i(1-p_i)
\end{equation}

where $n$ is the number of trials and $p$ is the probability. Here, I consider each row of a movement matrix a multinomial. This is what we discussed in the proposal defense. 
code modified from [\color{blue}here](http://stackoverflow.com/questions/19960605/r-multinomial-distribution-variance)

#### \color{red}Question: 
I'm not sure how to represent $n$, but I assumed the transition between seasons was a single trial for each row. 

<!-- L(x_{sst(t)}) = \int_{SST_{min}}^{SST_{max}} N(t; \mu_z, \sigma_z)~dz -->

```{r, echo = T, eval = T, warning = F, message = F, comment=F }
# function to calculate variance of a multinomial row
get.mvar = function(x){
  p = drop(x) # use the percentages
  v = p*(1-p)
  # c = outer(p, 1-p)
  # vcv = diag(c)
  v
}

# function to sample from a posterior multinomial distribution
get.mvar.samp = function(vmu, vsd) {
  o = rtruncnorm(1, mean = vmu, sd = vsd, a = 0, 1)
  o[is.nan(o)] = 0
  o
}

# example for seasonal transition matrices
vmat = lapply(watl.big, function(x) apply(x, 1, get.mvar))
```

```{r, vmattab, echo = F, eval = T, warning = F, message = F, comment=F  }
kable(vmat[[1]], row.names = 1:7, col.names = 1:7, caption = "Multinomial variance for Western bluefin >185cm, in Winter", digits = 4)
```

\newpage

## Plot the variance for each cell, each season (colorbar)
### Overlay with original movement matrix values

Example using Western bluefin >185cm

```{r, eval = T, echo = F, fig.width = 9, fig.height=9, warning = F, message = F, comment=F, results = 'hide'}
ex7 = expand.grid(1:7, 1:7)

par(mfrow=c(2,2))
sapply(1:4, function(x) {
  image.plot(1:7, 1:7, vmat[[x]], col = mycol(49), xlab = '', ylab = '')
  text(ex7[,1], ex7[,2], round(as.vector(watl.big[[x]]),3), font = 2)
  title(names(watl.big)[x])
  }
  )

```

----

\newpage

## sample from posterior
```{r, eval = T, echo = T, warning = F, message = F, comment=F}
samp = data.frame(sapply(1:7, function(x) get.mvar.samp(watl.big[[1]][x,], vmat[[1]][x,])))
names(samp) = 1:7
kable(samp, row.names = 1:7, caption = 'Posterior sample: Western Atlantic bluefin > 185cm, Winter season', digits = 4)
```

----

\newpage

## Goodness of Fit

### function: compare pairs of rows of two matrices, using loglinear model
returns the pchisq value

$\chi^{2}$ multinomial goodness of fit test from [\color{blue}here](http://www.r-tutor.com/elementary-statistics/goodness-fit/multinomial-goodness-fit)

\begin{equation}
{\displaystyle \operatorname {\chi^{2}} = \sum_{i} \frac{(f_{i}-e_{i})^2}{e_{i}}\,} 
\end{equation}

```{r, echo = T, eval = T, warning = F, message = F, comment=F}

# first arguement is counts, second argument is probability. In this way, row-wise comparisons of two matrices may be made
# p-values of the log-linear (chisq test) are returned

get.loglin <- function(mat1, mat2){
  fm = NULL
  for(i in 1:nrow(mat1)){
    fm[[i]] =  loglin(rbind(mat1[i,], mat2[i,]+1e-30),c(1), print = F, fit = F)
  }
  unlist(lapply(fm, function(x) 1-pchisq(x$lrt, x$df)))
}

get.mgof <- function(mat1, mat2){
  fm = NULL
  for(i in 1:nrow(mat1)){
    fm[i] =  chisq.test(mat1[i,], p = mat2[i,]+1e-30)$p.value
  }
 fm
}
```

I tried several versions of log-linear and $\chi^2$ tests. Log linear models, and their GLM versions, are essentially wrappers on $\chi^2$ tests for ease of use and flexibility in using contingency tables in various formats. 

Since each row of the movments matrices is dependent on it's associated region i nthe previous time step, the matrix as a whole cannot be considered a contingency table. Each row, however, may be considered it's own 1-dimensional contingency table.  

To compare any two rows, within a matrix or between matrices, the best method appeared to be the $\chi^2$ test. To set this up, counts within a row are compared to the probablity in another row. The probability from the second matirx row is simply $\frac{x_{i}}{\sum x_{i} }$, and is the markovian result we have used to date. 

To do goodness of fit comparisons, I used the `get.mgof` function. If any probabilities are $0$, the $\chi^2$ has issues. I added $1e-30$ to alleviate this problem. Perhaps there is a better way?

\newpage

### Compare between seasons of same run/stock (e.g. Western Atlantic > 185cm)

```{r, echo = T, eval = T, warning = F, message = F, comment=F}
bw.bs = get.mgof(watl.big.ct[[1]], watl.big[[2]])
print(bw.bs)
```


### Compare between runs, same season

```{r, echo = T, eval = T, warning = F, message = F, comment=F, results = 'asis'}
# same row, same season, large and small WABFT
# (bg.sm = get.loglin(watl.big.ct[[1]], watl.small[[1]]))

bg.sm = data.frame(sapply(1:4, function(x) get.mgof(watl.big.ct[[x]], watl.small[[x]])))
names(bg.sm) =  names(watl.big)
kable(bg.sm, row.names = 1:7, caption = "Western Atlantic bluefin > 185 cm vs < 185 cm", digits = 4)
```

\newpage

### Compare stocks and season
```{r, echo = T, eval = T, warning = F, message = F, comment=F, results = 'asis'}
e.w.s = data.frame(sapply(1:4, function(x) get.mgof(watl.big.ct[[x]], eatl[[x]])))
names(e.w.s) =  names(watl.big)

# library(xtable)

# print(xtable(e.w.s, include.rownames = T), type = 'html')
kable(e.w.s, row.names = 1:7, caption = "Western Atlantic bluefin > 185 cm vs Eastern Atlantic bluefin", digits = 4)
```

### Compare between stocks, non-seasonal
#### this doesn't work... need a different test if this comparison is neccesary

```{r, echo = T, eval = F, warning = F, message = F, comment=F}

# get overall transition rate; multiply the seasonal matrices
b1 = watl.big
b2 = eatl

ball.w = b1[[1]]%*%b1[[2]]%*%b1[[3]]%*%b1[[4]]
ball.e = b2[[1]]%*%b2[[2]]%*%b2[[3]]%*%b2[[4]]

(e.w = get.loglin(ball.e, ball.w))
```

----
\newpage

### Tradeoff between degrees of freedom and likelihood ratio within the $\chi^{2}$ test
this helped answer a question I had earlier
<!-- it seems the $\chi^{2}$ p-value is 1-pchisq.. when I do this, almost nothing is significantly different between matrices...   -->


```{r, eval = T, echo = F}
plot(pchisq(rep(50,100), df = 1:100), typ = 'l', lwd= 2, col = 4, xlab = 'DF', ylab = 'Value')
title('pchisq value')
```

