---
title: "Simulate using SatTagSim"
author: "Benjamin Galuardi"
date: '`r Sys.Date()`'
output:
  pdf_document:
    fig_caption: yes
  html_document: null
---
<!-- rmarkdown::html_vignette: -->
<!--     fig_caption: yes -->
<!-- vignette: > -->
<!--   %\VignetteIndexEntry{Simulate using SatTagSim} -->
<!--   %\VignetteEngine{knitr::rmarkdown} -->
<!--   %\VignetteEncoding{UTF-8} -->

##Simulation of Nova Scotia tagged Atlantic bluefin tuna \emph{(Thunnus thynnus)} and get seasonal transition matrices from a 7-box spatial strata.


```{r, global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, 
											message = FALSE, cache = FALSE,
											progress = TRUE, verbose = FALSE, fig.width = 8, fig.height = 6)

```	

load background data and set up parallel

```{r, setup_data, eval = T}
library(SatTagSim)
library(analyzepsat)
library(rworldmap)
library(rworldxtra)
library(knitr)
data(box7)
data(box11)
data(rmask)
data(nsfish)
data(myramps)
map = rworldmap::getMap(resolution = 'high')
# sst = rmask2array(rmask)
# mcoptions = setup.parallel()
```

## Tag data:
```{r, plot_tags, fig.width = 9, fig.height = 6}
par(mar = c(3,3,3,6))
plot(nsfish, col = month.colors[nsfish$Month,2], pch = 19, cex = .6, ylim = c(20, 50), xlim = c(-100, 0))
plot(map, add=T, col = 'cornsilk4', fill=T)
degAxis(1)
degAxis(2)
box()
 .add.month.scale()
```

\newpage


```{r}
# length of deployments
ddply(as.data.frame(nsfish), c('TagID'), function(x) data.frame(nmonths = length(unique(x$Month))))
```


\newpage

## plot as raster:
```{r, plot_tags_raster, fig.width = 8, fig.height = 6}

make.tag.raster  <- function(simdat, xmn = -100, xmx = 30, ymn = 0, ymx = 60, boxsize= 60){
  # require(raster)
  if(class(simdat)=="SpatialPointsDataFrame"){
      simdat = as.data.frame(simdat)
  }
  r = raster(nrow=(ymx-ymn)*60/boxsize, ncol= (xmx-xmn)*60/boxsize, xmn = xmn, xmx = xmx, ymn = ymn, ymx = ymx)
  simdat$CID = 1
  coordinates(simdat) = ~Longitude+Latitude
  sr = rasterize(simdat, r, field = 'CID', fun='count') #[[1]]
  return(sr)
}

r = dlply(as.data.frame(nsfish), 'Month', function(x) make.tag.raster(x))

r = make.tag.raster(nsfish)

library(viridis)
plot(r, col = rev(viridis(100)), ylim = c(20, 50), xlim = c(-100, 0), axes =F)
world(add=T, col = 'cornsilk4', fill=T)
degAxis(1)
degAxis(2)
box()

```

<!-- \newpage -->

<!-- ## Length of deployment by size -->
```{r, plot_DAL, eval = F}
# pdat = ddply(as.data.frame(nsfish), .(TagID, Length), .fun=function(x) data.frame(DAL = nrow(x)))
# 
# ggplot(pdat, aes(TagID, DAL))+geom_bar(stat = 'identity')
# 
# ggplot(as.data.frame(nsfish), aes(Longitude, Latitude, group = TagID, colour = MaxTemp))+geom_point(stat = 'identity')+coord_map()+scale_color_continuous(low = 'lightblue', high = 'red4', na.value = 'white')

```


\newpage
## Spatial strata in this package
7-box strata from the Kerr et al. operational model

```{r, plot_box7, fig.show = 'asis', echo = F, eval = T, fig.width = 8, fig.height = 6, fig.cap = '7- box stratification from Kerr et al. (2016) operational model'}
# library(fields)
# library(raster)
# library(SatTagSim)
# data(box7)
plot(box7, xlim = c(-100, 40), ylim = c(0,65))
world(add=T)
plot(box7, add = T, border = 'salmon', lwd=2)
text(coordinates(box7)[,1], coordinates(box7)[,2], box7@data[,3], cex = 2, col = 'darkred')
degAxis(1)
degAxis(2)
box()
title('7-box')
```

\newpage

```{r, plot_box11, fig.show = 'asis', echo = F, eval = T, fig.width = 8, fig.height = 6, fig.cap = '11- box stratification from Lauretta et al. (2016) spatial summary'}

plot(box11, xlim = c(-100, 40), ylim = c(-50,80))
world(add=T)
plot(box11, add = T, border = 'salmon', lwd=2)

lpts = coordinates(box11)
lpts[4,2] = 35
lpts[9,] = c(-25, 25)

text(lpts[,1], lpts[,2], box11@data[,3], cex = 2, col = 'darkred')
degAxis(1)
degAxis(2)
box()
title('11-box')
```


\newpage

## Environmental data in this package
World Ocean Atlas sea surface temperature is included. A raster mask, predicated on suitable temperature habitat in each month, is also included. 

```{r, plot_WOA, eval = T, fig.show = 'asis', fig.cap = 'World Ocean Atlas (2013) mean temperature'}
library(fields)
library(raster)
library(matlab)
data(rmask)
data(woasst)

# par(mfrow=c(1,2))
par(mar = c(2,2,2,4))
image.plot(woasst$lon, woasst$lat, apply(woasst$sst, 1:2, mean, na.rm = T), axes = F)
map(add = T, col = 'white', fill = T)
# world(add=T) 
degAxis(1)
degAxis(2)
box()
title('World Ocean Atlas mean temperature')

# image.plot(sstmat$lon, sstmat$lat, apply(sstmat$data, 1:2, sum, na.rm=T), xlim = c(-100, 40), ylim = c(0,65))
# world(add=T, col = 'white')
# title('Number of months with suitable surface temperature')
```


\newpage

## Set up the temperature constraint field. 

This can be done several ways. The preferred method is to use a running average of suitability, by month, for 12 months. In this manner, there is not a hard limit as to whether a fish may be in the area or not, but does provide a selection criteria that is weighted toward the suitability in that month.

The temperature field provided `data(rmask)` is a surface temperature suitability for bluefin tuna, based on tag measured temperatures. 

```{r, plot_ecdf, eval = T, fig.show = 'asis',  fig.cap = 'Tag Measured Temperatures'}
dat = nsfish

par(mfrow=c(3,4))
for(i in 1:12){
	plot(ecdf(dat$MaxTemp[dat$Month==i]), main = month.name[i], xlab = '' ,ylab = '')
	# hist(dat$MaxTemp[seasidx[[i]]], breaks=30, main=seasons[i], xlim = c(0, 40))
	mtemp = mean(dat$MaxTemp[dat$Month==i], na.rm=T)
	sdtemp = sd(dat$MaxTemp[dat$Month==i], na.rm=T)
	abline(v=mtemp, lwd=2, col=2)
	abline(v=quantile(dat$MaxTemp[dat$Month==i], c(0.025, .975),na.rm=T), lwd=2, col=4)
	# print(quantile(dat$MaxTemp[dat$Month==i], c(0.025, .975),na.rm=T))
}
```
Based on the plots, a range of $10-28\,^{\circ}\mathrm{C}$ were chosen 

\newpage

```{r, sst_pref_plot, fig.cap = 'Number of months with suitable surface temperature'}
sstmat = list(lon = sort(unique(coordinates(rmask[[1]])[,1]))
                ,lat = sort(unique(coordinates(rmask[[1]])[,2]))
                # ,data = as.array(rmask)
                )
  sstdata = array(NA, dim = dim(rmask)[c(2,1,3)])
  for(i in 1:(dim(rmask)[3])) sstdata[,,i] = rot90(as.array(rmask)[,,i],3)
  
  sstdata[is.na(sstdata)] = 0
  

# sstmat$data = apply(sstdata, 1:2, sum, na.rm = T)

# sstdata[is.na(sstdata)] = 0
sstmat$data = sstdata

image.plot(sstmat$lon, sstmat$lat, apply(sstmat$data, 1:2, sum, na.rm=T), xlim = c(-100, 40), ylim = c(0,65), col = viridis(12), axes = F)
plot(map, add=T, col = 'white')
# world(add=T, col = 'white')
degAxis(1)
degAxis(2)
title('Number of months with suitable surface temperature')

# rm(sstdata, sstdata2)

```


```{r, sst_pref_plot_12months, fig.cap = 'Number of months with suitable surface temperature. This plot shows a rolling sum of the current and next months binary values. This assists the simulation avoid entrainment in poor areas while accounting for where the fish might go next.'}
# 
# sstdata2 = sstdata
#   for(i in 1:12){
#     idx = c(i-1, i, i+1)
#     if(i == 1){ 
#         idx = c(12, 1, 2)
#     }
#     if(i == 12){
#       idx = c(11, 12, 1) 
#     }
#     # else idx = c(i-1, i, i+1)
#     sstdata2[,,i] = apply(sstdata[,,idx], 1:2, sum, na.rm = T)
#   }

sstdata[sstdata==0] = NA

sstdata2 = sstdata

  for(i in 1:12){
    idx = c(i, i+1)
    if(i == 1){
        idx = c(12, 1, 2)
    }
    if(i == 12){
      idx = c(11, 12, 1) 
    }
    else idx = c(i-1, i, i+1)
    sstdata2[,,i] = apply(sstdata[,,idx], 1:2, sum, na.rm = T)+sstdata[,,i]
  }

sstdata2[is.na(sstdata2)] = 0  
sstmat$data = sstdata2

par(mfrow=c(3,4))
for(i in 1:12){
  par(mar=c(1,1,2,5))
plot(map, add=F, col = 'white', xlim = c(-100, 40), ylim = c(0,65), axes = F)
image(sstmat$lon, sstmat$lat, sstmat$data[,,i], xlim = c(-100, 40), ylim = c(0,65),  col = viridis(max(sstmat$data[,,i])+1), axes = F, add=T)
plot(map, add=T, col = 'white', axes = F)
image.plot(sstmat$data[,,i], legend.only = T, col = viridis(max(sstmat$data[,,i])+1), nlevel = 3)
box()
# world(add=T, col = 'white')
degAxis(1)
degAxis(2)
# degAxis(1, tck = 0.02, las = 2, hadj = -1)
# degAxis(2, tck = 0.02, las = 1, hadj = -1)
title(month.name[i])
}
# title('Number of months with suitable surface temperature')

```

\newpage

```{r, plot_sstdat, eval = F, fig.show = 'asis'}
# SET UP THE WOA SST AS A 3D ARRAY
# Function does not work!!!
# sstmat = rmask2array(rmask)
sstmat = list(lon = sort(unique(coordinates(rmask[[1]])[,1]))
                ,lat = sort(unique(coordinates(rmask[[1]])[,2]))
                # ,data = as.array(rmask)
                )
  sstdata = array(NA, dim = dim(rmask)[c(2,1,3)])
  for(i in 1:(dim(rmask)[3])) sstdata[,,i] = rot90(as.array(rmask)[,,i],3)
  # sstmat$data = apply(sstdata, 1:2, sum, na.rm = T)

sstmat$data = sstdata
## If using the running average suitability
{
  sstm = sstdata
  for(i in 2:11){
   sstm[,,i] = apply(sstdata[,,c(i-1,i,i+1)], 1:2, sum, na.rm = T)
  }
  sstm[,,1] = apply(sstdata[,,c(12, 1, 2)], 1:2, sum, na.rm = T)
  sstm[,,12] = apply(sstdata[,,c(11, 12, 1)], 1:2, sum, na.rm = T)

  sstmat$data = sstm
  rm(sstdata, sstm)
}


par(mfrow=c(3,4))
for(i in 1:12){
  par(mar=c(2,2,2,2))
 image(sstmat$lon, sstmat$lat, sstmat$data[,,i], col = tim.colors(4))
 world(add=T)
}
```

## Get parameters from tag data

```{r, get_params, echo = T}
par.ns = get.allpar(as.data.frame(nsfish))
d.ns = get.kfD(as.data.frame(nsfish))
simpar = merge.par(par.ns, d.ns, return.mean = T)
```

```{r, show_params_table}
knitr::kable(simpar, caption = 'Advection and Diffusion parameters for simulation. For these fish, a fixed diffusion was used during track estimation so there is no variance in D')
```

\newpage

## Examine the tag based Advection and Diffusion parameters
```{r, D_plot, fig.cap = 'Distribution of diffusion (D) estimates by length (estimated CFL). All values here are identical as the track estimation routine used a fixed diffusion parameter. Here, D is measured as nautical $miles^2$/day'}

pdat = ddply(as.data.frame(nsfish), .(TagID), .fun=function(x) c(x$Length[1], x$D[1], x$Dsd[1]))
names(pdat)[2:4] = c('CFL','D','Dsd')

ggplot(pdat, aes(x=CFL, y=D)) + 
    geom_errorbar(aes(ymin=D-Dsd, ymax=D+Dsd), colour="black", width=.1) +
	geom_abline(aes(intercept=mean(D, na.rm=T), slope=0, width = 2),  colour='red')+
	# geom_abline(aes(intercept=mean(D+Dsd, na.rm=T), slope=0, colour='blue', width = 2))+
	# geom_abline(aes(intercept=mean(D-Dsd, na.rm=T), slope=0, colour='blue', width = 2))+
    geom_point( size=3, shape=21, fill="white") + # 21 is filled circle
	# theme(axis.text.x  = element_blank())+
    ggtitle(paste0("D distributions for ",nrow(pdat) ," tagged fish") )
```

\newpage

```{r, uv_plot, fig.cap = 'Distribution of monthly advection parameters. Wider distributions indicate greater variance. Units are nautical miles/day'}
plot.uv.density(par.ns)
```

\newpage

## Setup simulation parameters

```{r, set_simpar, echo = F, eval = T}

#------------------------------------------------------------------------#
# Set number of sims, simulation steps per month and number of years

msims = 50  # simulations to be started in each month
npmon = 30 # number of simulation steps per month.
nyears = 2 # number of years to simulate for each track
sstol = 2 # number of suitable months for each daily simulation step
#------------------------------------------------------------------------#

simvals = data.frame(sim_param = c('msims','npmon','nyears','sstol'), value = c(msims, npmon, nyears, sstol), description = c('simulations to be started in each month','number of simulation steps per month','number of years to simulate for each track','number of suitable months for each daily simulation step'))

kable(simvals, caption = "Control parameters for simulation")

#------------------------------------------------------------------------#
# Set month order of simulaitons to correspond to release month
morder = array(rep(1:12, 12), dim=c(12, 12))

for(i in 1:12){
  if(i==1) morder[i,] = 1:12
  else morder[i,] = c(i:12,1:((1:12)[i]-1))
}
#------------------------------------------------------------------------#

morder = data.frame(morder, row.names = month.abb)
names(morder) = paste0('m',1:12)
kable(morder, caption = 'Month order for multi-month start simulation', row.names = T)
```

\newpage

## Set up starting points for multi-start month simulation

Set any indices for size or other criteria here. 

```{r, stat_points, eval = T, echo = T}
ns = as.data.frame(nsfish)
ns = ns[,c('TagID','Day','Month','Year','Longitude','Latitude')]
spts = get.start.pts(ns, msims, months = 1:12, posnames = c('Longitude','Latitude'))
str(spts)
# plot(map, xlim = c(-100, 0), ylim = c(20, 50))
# for(i in 1:12) points(spts[[i]], col = month.colors[which(as.numeric(month.colors[,1])==i),2], pch = 19)
# degAxis(1)
# degAxis(2)
# box()

```

\newpage

## Run a single simulation in parallel

```{r, simulate, eval = T}

mcoptions = setup.parallel()

#---------------------------------------------------------------#
# SIMULATIONS

print(paste0('simulating ', length(spts)*msims, ' tracks for ', nyears, ' years'))
stime = Sys.time()
simdat = list()

ncores = detectCores()
cl = makeCluster(ncores)
registerDoParallel(cl, cores = ncores)

for(i in 1:12){
  subsp = spts[[i]][sample(1:nrow(spts[[i]]), msims, replace = T),]
  subsp$row = 1:nrow(subsp)
  sp = dlply(subsp, 'row', function(x) x[,1:2])
  simorder = as.numeric(morder[i,])
  
  print(paste0('simulating ', length(sp), ' tracks starting in ', month.name[i]))
  
  test = make.sim.track.par(tpar = simpar, morder = rep(simorder, nyears), sp = sp, bath = bath, sstmat = sstmat, seaslen = npmon, sstol = sstol, mcoptions = mcoptions)
  simdat[[i]] = test
  runtime = Sys.time()-stime
  print(paste0('elapsed time: ', runtime))
}

stopCluster(cl)
rm(test, sp, subsp)

simdat = unlist(simdat, recursive = F)
```

\newpage

## Make a dataframe and raster of results

```{r, simdatdf, eval = T, fig.cap = 'Posterior distribution of simulated tracks; 1-minute cell size'}
# MAKE A DATA FRAME OF THE RESULTS
simdatdf = ldply(simdat)
seasons = c("Winter", "Spring", "Summer", "Fall")
# seasidx = make.seas.idx(simdatdf)
simdatdf$seas[simdatdf$Month%in%c(1,2,3)] = 1
simdatdf$seas[simdatdf$Month%in%c(4,5,6)] = 2
simdatdf$seas[simdatdf$Month%in%c(7,8,9)] = 3
simdatdf$seas[simdatdf$Month%in%c(10,11,12)] = 4
# rm(seasidx)

names(simdatdf)[which(names(simdatdf)%in%c('lon,','lat'))] = c('lon','lat')

#---------------------------------------------------------------#
# Make Rasters of Results

mycol = colorRampPalette(c("lightcyan", "royalblue", "blue", "lemonchiffon", "orange", "red"), space = "Lab")

sr = dlply(simdatdf, 'seas', function(x) make.sim.raster(x, boxsize = 60))

par(mfrow=c(2,2))
par(mar=c(2,2,2,2))
for(i in 1:4){
  plot(sr[[i]]/max(values(sr[[i]]), na.rm=T), zlim = c(0.01,1), interp=F, col=mycol(256), axes=F)
  rpts = xyFromCell(sr[[i]], which(!is.na(values(sr[[i]]))))
  world(add=T, col='grey80', fill=F)
  #   plot(box7, add=T, border='salmon', lwd=2)
  #   text(coordinates(box7)[,1], coordinates(box7)[,2], box7@data$ID, font=2, cex=1.2)
  plot(box11, add=T, border='salmon', lwd=2)
  # text(coordinates(box11)[,1], coordinates(box11)[,2], box11@data$ID, font=2, cex=1.2)
  text(lpts[,1], lpts[,2], box11@data[,3], cex = 2, col = 'black')
  degAxis(1)
  degAxis(2)
  title(seasons[i])
  box()
}

```
\newpage

## Get seasonal transition matrices using the 11 box model

```{r, get_transmat, eval = T}

# try new version of function... it works!!
datbox = get.first.box(simdat, 2000, box11)
boxtrans = get.trans.prob(datbox, nyears=100, adims = c(11,11,4), perc = F)

names(boxtrans) = seasons

plot.boxtrans(boxtrans)
```

